### 2016-02-02 manuscript_methods/Open_experiment.Rmd
Solidified outline, made ms document

---
title: "Running an open experiment: transparency and reproducibility in soil and ecosystem science"
author: "Ben Bond-Lamberty"
date: "February 2, 2016"
output: html_document
---

#### Introduction

Look at - Ruegg et al. paper on data life cycle!
Data paper - https://peerj.com/preprints/1448/
http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.1002295

Peyton's notes:

>Add to Introduction: 
>Reichman et al 2011 Science 331.pdf
>
>Wolkovich, E. M., Regetz, J. and O'Connor, M. I. (2012), Advances in global change research require open science by individual researchers. Glob Change Biol, 18: 2102–2110. doi:10.1111/j.1365-2486.2012.02693.x
>
>Add to reproducibility literature:
>http://journals.plos.org/plosbiology/article?id=10.1371%2Fjournal.pbio.1002333

#### Why more openness in science?

* Reproducibility

Thornton paper on archiving models

* Data loss

17% per year paper - Vines et al.

* Data quality

"All bugs are shallow" Raymond quote

* Increasing transparency of the process - increase public trust and access to science

* Value of collected databases - makes this easier
* BAAD TRY, FLUXNET, Dong-Gill, SRDB

#### Push and pull: funders, journals, and society

PLOS ONE, ESA, etc. policies

* NSF url
* DOE url
* NASA url
* NSERC, Wellcome trust, other?

#### Running a fully open experiment: one example

* GitHub
* Diagnostics page(s)
* Data: raw to summarized to final
* All statistical analyses
* "LRB" - notes 
* Manuscript(s)

#### Conclusions

This model not appropriate for everyone/thing!

Potential concerns

* Not appropriate for every kind of study
* Fear of "being scooped"
* Question of open from beginning or only after publication
* Privacy concerns