### 2016-02-02 manuscript_methods/Open_experiment.Rmd
Refining outline, adding citations and thoughts

---
title: "Running an open experiment: transparency and reproducibility in soil and ecosystem science"
author: "Ben Bond-Lamberty, Peyton Smith, and Vanessa Bailey"
date: "February 2, 2016"
output: html_document
---

#### Introduction

General intro.

PLOS ONE, ESA, etc. policies

* NSF url
* DOE url
* NASA url
* NSERC, Wellcome trust, other?

*Reproducibility*

Surveying the 2000-2014 biomedical literature, Iqbal et al. {, 2016 #4334} found that none of 441 randomly-chosen studies provided raw data, and only one provided full protocols.

"Reproducibility of experimental results is a fundamental criterion for credible scientific research" {Thornton, 2005 #2310}. More difficult for modeling work; this is a subset of the more general problem with respect to digital media, software, etc. (CITATION?).

Scripted systems for data analysis! E.g. R {R Development Core Team, 2016 #4314}, Kepler (https://kepler-project.org) or Taverna (http://www.taverna.org.uk) record all infrmation about the data processing, analytical process and decisions, and statistical analysis.

*Data loss*

Vines et al. {, 2014 #4130} published a shocking result, based on surveying 516 articles from 2 to 22 years old: the odds of a data set being available fell by 17% each year, and the chances that the contact author's email address still worked declined by 7% per year. Reichman et al. {, 2011 #2862} estimated that less than 1% of ecological data collected is made available after publication, and noted, as an example, that most current and historical data relevant to the recent *Deepwater Horizon* oil spill are already inaccessible or lost {Reichman, 2011 #2862}.

Scientists' use of strong and consistent data curation practices {Hart, 2015 #4292}{Rüegg, 2014 #3886} can mitigate the problem, but all our anecdotal experience and quantitative studies {Vines, 2014 #4130} suggest that in the long term, data cannot be reliably preserved by individual researchers.

Data loss hits ecosystem, soil, and global change ecology particularly hard, as global to local climate changes make ecological data effectively irreproducible {Wolkovich, 2012 #3303}: we can never remeasure exactly the same system state.

*Data and code quality*

Providing open code does not magically produce bug-free code, or mistake free analyses {Easterbrook, 2014 #4137}. But it does encourage authors to invest the time upfront to clean up their code, data, and documentation when a paper is written, rather than deferring this task (often, forever).

Potential to increase pulic trust in research results.

*Enabling synthetic and global-scale analyses*

Many refs in {Wolkovich, 2012 #3303}.

BAAD {Falster, 2015 #4209}, TRY {Kattge, 2011 #3064}, FLUXNET {Baldocchi, 2001 #2883}, SRDB {Bond-Lamberty, 2010 #2320}.

#### An open experiment: one example

This is example, we hope, of the impact of individual scientists {Wolkovich, 2012 #3303}.

* GitHub
* Diagnostics page(s)
* Data: raw to summarized to final
* All statistical analyses
* "LRB" - notes 
* Manuscript(s)

#### Conclusions

Do we adequately credit researchers who contribute to global databases {Kattge, 2011 #3064} that subsequently produce high-impact papers {Díaz, 2016 #4318}?

Such meta-analyses rely on the collection of primary data, and it is critical that field and experimental researchers efforts are adequately valued and cited {Kueffer, 2011 #4335}.

This model not appropriate for everyone/thing!

Potential concerns

* Not appropriate for every kind of study
* Fear of "being scooped"
* Question of open from beginning or only after publication
* Privacy concerns