### 2016-06-28 manuscript_openexperiment/Response_to_reviewers.Rmd
Added response to reviewers doc.

---
title: "Response_to_reviewers"
author: "Ben Bond-Lamberty"
date: "June 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```